---
title: "proposal"
author: "Tristan Shah"
date: "2/16/2022"
output: html_document
---

```{r}
library(tidyverse)
```

## Introduction: What is your research question? Why do you care? Why should others care? If you know of any other related work done by others, please include a brief description.

## Data: Include context about the data covering:

## Data source: Include the citation for your data, and provide a link to the source.

## Data collection: Context on how the data was collected?

## Cases: What are the cases (units of observation or experiment)? What do the rows represent in your dataset?

## Variables: What are the variables you will be studying?

## Type of study: was it an observational study or an experiment?

## Data Quality: Check for data quality issues, missing data, duplicates, format issues etc. and perform the necessary quality improvements. 

## References: Any write-ups or analysis you might have referenced for this project.

```{r}
tibble(read.csv("san-jose-fire-incidents-data-2020.csv")) -> fires
tibble(read.csv("fire_station_locations.csv")) -> stations
```


```{r}
## Dropping all incomplete rows
fires <- fires[complete.cases(fires), ]

## Obtaining vector of unique stations in the locations dataset
valid_stations <- unique(stations$STATION)

## Filtering out stations with irregular names.
## Also only selecting stations which are present in the locations dataset
fires %>%
  filter(nchar(Station) == 2) %>%
  mutate(Station = as.integer(Station)) %>%
  filter(Station %in% valid_stations) -> fires

## Cleaning the Priority data to be of type int
fires$Priority %>% 
  strsplit(" ") %>% 
  map_chr(2) %>% 
  as.integer() -> fires$Priority

## Cleaning lat and long
fires$fire_long <- as.numeric(fires$Near_X)
fires$fire_lat <- as.numeric(fires$Near_Y)

## Selecting columns of interest
fires[, c(
  "fire_long", 
  "fire_lat", 
  "Priority", 
  "Date_Time_Of_Event",
  "Dispatched_Time",
  "Unit_On_The_Way_Time",
  "Cleared_TimeStamp",
  "Final_Incident_Type",
  "Final_Incident_Category")] -> fires

## Creating logical vector which determines if any time column is missing a value
fires$Date_Time_Of_Event != "" &
  fires$Dispatched_Time != "" &
  fires$Unit_On_The_Way_Time != "" &
  fires$Cleared_TimeStamp != "" -> valid_times

## Removing rows with missing times
fires[valid_times,] -> fires

head(fires)
```

```{r}
convert_timestamp <- function(time_column) {
  return(strptime(time_column, format = "%m/%d/%Y %R %p"))
}

mutate(fires,
  across(
    c(
      Date_Time_Of_Event, 
      Dispatched_Time, 
      Unit_On_The_Way_Time, 
      Cleared_TimeStamp), convert_timestamp)) -> fires

time_to_dispatch <- fires$Dispatched_Time - fires$Date_Time_Of_Event
dispatch_to_leave <- fires$Unit_On_The_Way_Time - fires$Dispatched_Time
leave_to_cleared <- fires$Cleared_TimeStamp - fires$Unit_On_The_Way_Time

sum(delta < 0, na.rm = T)
```
